PANCHANGE_data_preprocessing

### 2020-08-13 PANCHANGE_data_preprocessing2020-07-15.Rmd
Added chunk for RAMP into highest education.

## Highest education (baseline)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education baseline}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

@@

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```

>>>>

## RAMP
### Read in data
```{r ramp employment cols}
if(RAMP == TRUE){
ramp.employment.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_coping_ramp.rds"))
dim(ramp.employment.raw)
colnames(ramp.employment.raw)
}
```

### Select variables
```{r ramp employment cleaning}
if(RAMP == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

ramp.employment.raw.id <- ramp.employment.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(ramp.employment.raw.id)
# Inspect colnames
colnames(ramp.employment.raw.id)
#Differences
dim(ramp.employment.raw)[1]-dim(ramp.employment.raw.id)[1]
}
```

```{r ramp copy employment}
if(RAMP == TRUE){
  employment.raw.id <- ramp.employment.raw.id
}
```

## Highest education (baseline)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education baseline}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP = TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        employment.phd_numeric == "1" ~ 4,
        employment.masters_degree_or_equivalent_numeric == "1" ~ 4,
        employment.postgraduate_degree_or_equivalent_numeric == "1" ~ 4,
        employment.bachelors_degree_or_equivalent_numeric == "1" ~ 4,
        employment.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        employment.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        employment.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        employment.cses_or_equivalent_numeric == "1" ~ 1
      )
    )

@@

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP = TRUE) {
employment.raw.id %>%
  select(highest_education,
         employment.phd_numeric,
         employment.masters_degree_or_equivalent_numeric,
         employment.postgraduate_degree_or_equivalent_numeric,
         employment.bachelors_degree_or_equivalent_numeric,
         employment.a_levelsas_levels_or_equivalent_numeric,
         employment.nvq_or_hnd_or_hnc_or_equivalent_numeric,
         employment.o_levelsgcses_or_equivalent_numeric,
         employment.cses_or_equivalent_numeric
  )
}
```
